Contracts for unions and intersections are hard for Nickel

### 2021-05-18 paper.tex
some comments
In-call modifications

\begin{figure}
  \begin{center}
\begin{lstlisting}[language=nickel,title={Source program}]
let f = fun x => g y (x + 1) 
\end{lstlisting}
\begin{lstlisting}[language=nickel,title={Optimized program}]
let f' = let g' = g y in
  fun x => g' (x + 1)
\end{lstlisting}
  \end{center}
\caption{Let-floating}
\label{fig:optimizations-let-floating-ex}
\end{figure}

@@

However, this property does hold for \emph{pure} terms, that are terms without
side-effects. In a pure language, or a language with effect tracking like
Haskell, pure terms can be safely identified for applying optimizations. Other
languages rely on under-approximations.

\unsure{Should we tease already here that unions and intersections break this
even for pure terms?}
\newpage

@@

% Content:
%   - Unions are useful as a fundamental data block
%     Examples: nullable, tree-like data structure, multiple representations

\begin{lstlisting}
type BinaryTree =
Integer | (BinaryTree, BinaryTree)
\end{lstlisting}

\begin{lstlisting}
optionalValue : String | Null
\end{lstlisting}

\begin{lstlisting}
person : {name : String, age : Number}
  | {name : String, dateOfBirth : String}
\end{lstlisting}

%   - Intersections are useful for subtyping or refinement (integers/floats
%       e.g)
\begin{lstlisting}
1 : Int & Float
\end{lstlisting}
%   - Intersections are useful to encode overloading
\begin{lstlisting}
incr: (Int -> Int) & (Float -> Float)
\end{lstlisting}

%   - Intersections are useful to encode mixins (see Scala3)
\begin{lstlisting}
Comparator: {cmp: T -> T -> Bool}
Shower: {show: T -> String}

showSmaller : (dict: Comparator & Shower) -> T -> T -> String
\end{lstlisting}
%   - Intersections appears in some form as unions' dual
\begin{lstlisting}
\end{lstlisting}

% Things to settle: how to motivate them, how to compare them to ADTs. Purely
% dynamic approach, or typed approach too?
\newpage

@@

\begin{lstlisting}[language=racket]
(define/contract united
(or/c (-> number? number?)
      (-> even? even?))
(lambda (x) x))
\end{lstlisting}
\caption{Non working \racket{or/c} example on Racket.}
\label{code:racket:or/c:non-working}

>>>>

\begin{figure}
  \begin{center}
\begin{lstlisting}[language=nickel,title={Source program}]
let f = fun x => g y (x + 1)
\end{lstlisting}
\begin{lstlisting}[language=nickel,title={Optimized program}]
let g' = g y
let f' = fun x => g' (x + 1)
\end{lstlisting}
  \end{center}
\caption{Let-floating}
\label{fig:optimizations-let-floating-ex}
\end{figure}

@@

However, this property does hold for \emph{pure} terms, that are terms without
side-effects. In a pure language, or a language with effect tracking like
Haskell\todo{Cite a different language, or remove citation}, pure terms can be
safely identified for applying optimizations. Other languages rely on
under-approximations.

\subsubsection*{Lazyness}
\unsure{Should we talk about lazyness?
this would be a good place to introduce it}

\unsure{Should we tease already here that unions and intersections break this
even for pure terms?}

\subsection*{Laziness}

@@

Unions are useful as a fundamental data block.
Examples: nullable, tree-like data structure, multiple representations

\begin{lstlisting}[title={Tree-like data structure}]
type BinaryTree =
Integer | (BinaryTree, BinaryTree)
\end{lstlisting}

\begin{lstlisting}[title={Plain enumeration}]
protocol : `Http | `Ftp | `Sftp
\end{lstlisting}

\begin{lstlisting}[title={Nullable values}]
optionalValue : String | Null
\end{lstlisting}

\begin{lstlisting}[title={Alternative representations}]
person : {name : String, age : Int}
  | {name : String, dateOfBirth : String}
\end{lstlisting}

%   - Intersections are useful for subtyping or refinement (integers/floats
%       e.g)
\begin{lstlisting}[title={Value subtyping}]
1 : Int & Float
\end{lstlisting}
%   - Intersections are useful to encode overloading
\begin{lstlisting}[title={Overloading}]
incr: (Int -> Int) & (Float -> Float) =
  fun x => x + 1
\end{lstlisting}

%   - Intersections are useful to encode mixins (see Scala3)
\begin{lstlisting}[title={Structural mixins}]
Comparator: {cmp: T -> T -> Bool}
Shower: {show: T -> String}

showSmaller : (dict: Comparator & Shower) -> T -> T -> String
\end{lstlisting}
%   - Intersections appears in some form as unions' dual
\begin{lstlisting}[title={Intersection as the dual of unions}]
f : (Num | String) -> Bool
(*@ $\simeq$ @*) f : (Num -> Bool) (*@ $\cup$ @*) (String -> Bool)
\end{lstlisting}

\todo{Either here or the next section, show examples of unintuitive behaviour
of U and I. For example, I is not jos composition of contracts.
And union has to be treated as any one of the options}
% Things to settle: how to motivate them, how to compare them to ADTs. Purely
% dynamic approach, or typed approach too?

@@

\begin{lstlisting}[language=lisp]
(define/contract united
(or/c (-> number? number?)
      (-> even? even?))
(lambda (x) x))
\end{lstlisting}
\caption{Non working \racket{or/c} example on Racket.}
\label{code:racket:or/c:non-working}
